Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-16 manuscript.tex
discussion

As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate. A notable exception of a SNN using precise spike timings is the \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}. This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}. It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}. Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma. It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation. Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs). One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time. This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevich}). According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group. %----------------------------%

@@

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we have seen that the neurobiological system uses processing loops within hierarchical graphs. In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}. The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies. Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts).
%
\subsection{Perspectives}
%
The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.

>>>>

As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate. A notable exception of a SNN using precise spike timings is the \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}. This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}. It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}. Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma. It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation. Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs). One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time. This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevich}). According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group. 

@@

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we have seen that the neurobiological system uses processing loops within hierarchical graphs. In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}. The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies. Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts)~\citep{coull_distinction_2022}.
%
\subsection{Perspectives}
%
The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.
